|

### 2020-11-16 paper/v1/broadband.speed.covid.Rmd

**PARA2**The capability to work from home has previously been studied from the perspective of whether work tasks in a given occupation both can be and are allowed to be performed independently of location or co-location with colleagues, including supervisors (**insert refs**). However, successful telecommuting also requires that the quality and reliability of ICT services, particularly home internet connection speeds, enable the completion of work tasks with a minimum of delay or interruption. In reality, the performance of broadband speeds with respect to telecommuters has never been tested before at scale, as working from home and connecting to colleagues and workplace resources via broadband has previously been the purview of a small minority of workers. In the UK, the most recent statistics indicate that **x%** of working age adults identified their home as their usual workplace, with an additional **x%** who said they work from home at least once a week (**insert stats from 2018 or 2019 NTS - also x% of these use ICT to complete their work at home**). Their demands on internet services have been minor compared to the demands of leisure users, such that the broadband performance offered by different Internet Service Providers (ISPs) has been benchmarked according to download speeds during the evening 'primetime', when video streaming services are at their peak, rather than during the working  day (**OfCom, 2017**). Yet the pandemic has fundamentally changed not only how many people work from home - 47% of working age adults at the peak in April, 2020 (**ref**) - but also their technical requirements when doing so. Whilst it may have been common in the past for telecommuters to complete solitary work tasks at home, but still visit their workplace for meetings (**ref?**), during the pandemic the replacement of face-to-face contact with bandwidth-intensive video conference calling was seen as essential to suppressing the spread of infection. This type of broadband use can particularly affect upload speeds **ref?**, so we seek to develop a new benchmark (**is it a benchmark?**)to assess the quality and reliability of upload speeds by time of day and day of the week across local authorities in the UK during the time when the population were told to work from home if at all possible.

@@

>>>>

HB
**PARA2**The capability to work from home has previously been studied from the perspective of whether work tasks in a given occupation both can be and are allowed to be performed independently of location or co-location with colleagues, including supervisors (**insert refs**). However, successful telecommuting also requires that the quality and reliability of ICT services, particularly home internet connection speeds, enable the completion of work tasks with a minimum of delay or interruption. In reality, the performance of broadband speeds with respect to telecommuters has never been tested before at scale, as working from home and connecting to colleagues and workplace resources via broadband has previously been the purview of a small minority of workers. In the UK, the most recent statistics indicate that **x%** of working age adults identified their home as their usual workplace, with an additional **x%** who said they work from home at least once a week (**insert stats from 2018 or 2019 NTS - also x% of these use ICT to complete their work at home**). Their demands on internet services have been minor compared to the demands of leisure users, such that the broadband performance offered by different Internet Service Providers (ISPs) has been benchmarked according to download speeds during the evening 'primetime', when video streaming services are at their peak, rather than during the working  day (**OfCom, 2017**). Yet the pandemic has fundamentally changed not only how many people work from home - 47% of working age adults at the peak in April, 2020 (**ref**) - but also their technical requirements when doing so. Whilst it may have been common in the past for telecommuters to complete solitary work tasks at home, but still visit their workplace for meetings (**ref?**), during the pandemic the replacement of face-to-face contact with bandwidth-intensive video conference calling was seen as essential to suppressing the spread of infection. This type of broadband use can particularly affect upload speeds **ref?**, so we seek to develop a new benchmark (**is it a benchmark?**)to assess the quality and reliability of upload speeds by time of day and day of the week across local authorities in the UK during the time when the population were told to work from home if at all possible.

@@

ET
During the pandemic, working from home using Information and Communication
Technologies (ICT), whether partially or exclusively, was transformed from a
niche means of accessing work, albeit one that had been on a slow, upward trend,
to a widespread way of life in many countries.
The ability to work from home or telecommute meant millions retained their jobs
and, to a varying extent, maintained productivity during periods of strict
lockdown around the world.
However, this ability was not evenly distributed socially or spatially, creating
a new type of digital divide -- one of economic resilience and personal safety on
one side and unemployment or essential front-line work on the other.
The ability to telecommute is partly a product of whether the majority of work
tasks in an occupation can be performed independently of being in a particular
location or co-locating with colleagues.
It is also a product of the quality of ICT services, particularly home internet
connections, sufficient to successfully complete work tasks with a minimum of
delay or interruption.
Therefore, in this paper, we look at patterns of broadband speed tests
and upload speeds in the UK in order to gain insights into both the quality and
reliability of broadband services across the country and the distribution of
telecommuting at the time when the population were told to work from home if
possible.
To do so, we employ unique data regarding individual broadband speed tests and
state of the art time series clustering methods.
Our analysis enables us to better understand how telecommuting and
technology intersect at a time of extreme demand, and what lessons this time has
for a future where telecommuting is likely to remain a common means of accessing
work and broadband services, as well as infrastructure, must be fit for purpose.
**I THINK WE NEED A MORE CLEAR RESEARCH QUESTION**

<!-- How much do we work form home-->
The shift towards telecommuting during various stages of lockdown around the world
has been drastic and, importantly, speculations indicate that the post-Covid
tendency to work from home will be much higher than the pre-Covid one.
A back of the envelope calculation suggests that up to 40% of the working force
could work from home in the UK [@batty2020editorial].
Observational data pointed an even higher share of people in employment in the
UK who worked from home in April $2020$ ($47$%), while the same figure only reached
$5$% the year before [@ons2020; @ons2020lm2019].
Similar figures have been reported for other countries around the world
[@felstead2020homeworking].
For instance, $37$% of the workforce worked from home in Europe in April $2020$ with
countries like Finland reaching $60$% [@eurofound2020].
In the US almost half of the working population worked from home during the same
period because of the pandemic [@brynjolfsson2020covid].
More broadly, a recent estimate indicated that $37$% of all jobs in the US can be
permanently performed entirely from home [@NBERw26948].

<!-- Who works from home-->
There is a consensus that opportunities for working from home especially during
the current pandemic are not equally spread across the workforce.
@NBERw26948 indicated that in the US managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from
home.
On the contrary, workers in farming, construction, and manufacturing do not have
such opportunities.
Not surprisingly, occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in less footloose occupations as
as they tend to be lower-income, non-white, without a university degree,
live in rental accommodation and lack health insurance [@NBERw27085].
Although these figures refer to the US, similar trends can be observed for other
countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, while the same share for workers with secondary and
primary education dropped to $34$% and $14$% respectively.
Moreover, employees living in cities, women and younger employees were
have worked from home [@eurofound2020].

<!-- ICT infrastructure-->
None of these changes could have happened in the absence of ICT
infrastructure -- both in terms of software and hardware.
But while innovations in software are easily diffused across space and
society^[See for example the huge success of videoconferencing apps
[@marks2020zoom].],
the same does not apply for ICT hardware infrastructure such as internet
broadband connectivity.
The literature exemplifies digital divides in terms of internet access and its
quality.
For instance, @riddlesden2014broadband highlighted the broadband divides in the
UK, while the systematic review from @SALEMINK2017360 reinforced our
understanding for the infrastructure quality differences between urban and rural
areas.

<!-- WFH and digital divides-->
Our framework to understand telecommuting as a function of occupations and
quality of ICT infrastructure is aligned with current debates on digital divides.
While the, so-called, first level digital divides are associated with access
and quality of internet connectivity, the second level ones are linked
to the necessary skills to effectively utilise ICT and the internet
[@blank2014dimensions; @van2011internet].
Importantly, the differentiated capacity to telecommute, which to a certain
extend are related to the first and second level digital divides, leads to
differentiated outcomes regarding the economic resilience of people and
places against the current pandemic.
And to the extend that the quality of the internet infrastructure and occupation
variation are spatially dependent and clustered in space, the spatial footprint
of telecommuting is of great interest. In a way similar to the third level
digital divide, which
focuses on the differentiated returns of internet use [@stern2009levels;
@van2014digital; @van2015third] places with high rates of telecommuting during
the Covid pandemic illustrate higher economic resilience against the current pandemic.
<!-- Econ resilience-->
This is aligned with the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic].

<!-- WFH broader consequences-->
The long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they might span through various aspects of economy and society: from changes to
transportation planning due to changes in commuting patters to changes in land
use and urban planning to accommodate people who work from home
(@BUDNITZ2020102713); and from productivity and innovation changes to a change
in agglomeration externalities and the attraction of large cities [@econobs]
just to name a few. This paper is positioned to support such endeavours by
providing an explanatory **[NOTE SURE ABOUT THIS]** framework  based on
the quality of internet connectivity and the frequencies of occupations within
Local Authorities in the UK.

<!-- How can we observe observe WFM? -->
<!-- removed for now -->
<!-- Directly through surveys (@felstead2020homeworking).
Add survey limitations (cost, representativeness, time lag between survey and data availability)
Data gap.
Passive data collection through internet speeds
Another option could be mobility data.
How is internet speed data a better source than mobility?-->

**MORE SOURCES:**

- https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas
- [EPB commentaries](https://journals.sagepub.com/toc/EPB/current)
- https://www.coronavirusandtheeconomy.com/question/what-has-coronavirus-taught-us-about-working-home
- https://www.coronavirusandtheeconomy.com/question/who-can-work-home-and-how-does-it-affect-their-productivity
- https://www.coronavirusandtheeconomy.com/question/how-will-economic-effects-coronavirus-vary-across-areas-uk
- https://www.coronavirusandtheeconomy.com/which-parts-uk-have-been-hit-hardest-covid-19-crisis
- https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas

**PARA5**: Data and methods

**PARA6**: Contribution:

## broadband studies, divides, broadband tech stuff

## from telecommuting to #WFH
